Named Entity Recognition with Structured Prediction

### 2012-11-06 doc/report/paper.tex
Example layouts, some writing.

\documentclass[11pt]{article}
\usepackage{acl2011}
\usepackage[utf8]{inputenc}
\usepackage{linguex}
\usepackage{latexsym}  
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{array}
\usepackage{subfigure}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{qtree}

% \usepackage{xcolor}
% \definecolor{dark-red}{rgb}{0.4,0.15,0.15}
% \definecolor{dark-blue}{rgb}{0.15,0.15,0.4}
% \definecolor{medium-blue}{rgb}{0,0,0.5}
% \hypersetup{
%     colorlinks, linkcolor={dark-red},
%     citecolor={dark-blue}, urlcolor={medium-blue}}
\title{Named Entity Recognition using Structured Prediction}

@@

\section{Introduction}
Named Entity Recognition (NER) is an information extraction task and has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference)
that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
The most prominent types are generally known under the name of \emph{enamex} types, which are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}. 
An additional type \texttt{Miscellaneous} captures person names outside the classic \emph{enamax} type.
Apart from these there is \emph{timex}, which covers date \& time expressions and \emph{numex} which is for monetary values \& percent. 
Table \ref{table:NETypes} gives an overview of the different NE types.

\begin{table}[h!]
\scriptsize
\begin{tabular}{l l}
\bf Named Entity Type & \bf Example \\
\hline
Person Names & Greta Garbo, John, Mary \\
Organisations& Benetton, Coca-Cola Gmbh\\
Locations&  Paris, Australia, Bristol\\
Miscellaneous& World Cup\\
 Date Expression& 01-02-2012, January, 6th, 1998 \\
Time Expression & 10 p.m.\\
Monetary Value &  \$15, 100 Euro   \\
Percent &   30\% \\
\end{tabular}
\caption{NE Examples}
\label{table:NETypes}
\end{table}

@@

\ex. \textbf{x:} The motor company 'Ford' was founded and incorporated by Henry Ford.\\ \textbf{y:} O  O  ORG  O  O O O O PER PER.  \label{PredEx1b}

@@

Learning is implemented with a Structured Perceptron. To avoid overfitting, we average the parameters after the last iteration. For finding the best sequence of labels, we use a modified version of the Viterbi algorithm. The algorithm finds the sequence $\hat{y}$, such that:

@@

\paragraph*{Node features}
These are only present on the word in question: the \emph{Token}, suffix and prefix, capitalisation.

@@

\begin{table}[h!]
\begin{tabular}{l|l}
\hline
\bf Feature & \bf Example \\
\hline
Token &  \textbf{``Amsterdam''}\\
Suffix& Amster\textbf{dam}\\
Prefix&  \textbf{San} Sebastian\\
Captitalized& \textbf{B}enetton\\
Number Pattern & \\
UPPERCASE &  \textbf{BENETTON}\\
POS-tag &  Benetton \textbf{NNP}   \\
Lemma &  produced $Rightarrow$ produce \\
\end{tabular}
\caption{Node Features}
\label{table:node}
\end{table}

@@

\ex. Jack London went to New York. \\ PER   PER  0  0  LOC   LOC. \label{seq1}

\ex. Jack London went to New York. \\ PER   LOC  0  0  LOC   LOC. \label{seq2}

>>>>

\documentclass[11pt]{article}
\usepackage{acl2011}
\usepackage[utf8]{inputenc}
\usepackage{linguex}
\usepackage{latexsym}  
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{array}
\usepackage{subfigure}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{qtree}
%\usepackage{gb4e}

%\usepackage{gb4e}

% \usepackage{xcolor}
% \definecolor{dark-red}{rgb}{0.4,0.15,0.15}
% \definecolor{dark-blue}{rgb}{0.15,0.15,0.4}
% \definecolor{medium-blue}{rgb}{0,0,0.5}
% \hypersetup{
%     colorlinks, linkcolor={dark-red},
%     citecolor={dark-blue}, urlcolor={medium-blue}}
\title{Named Entity Recognition with Structured Prediction}

@@

\section{Introduction}
Named Entity Recognition (NER) is an information extraction task and has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference)
that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
The most prominent types are generally known under the name of \emph{enamex} types, which are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}. 
An additional type \texttt{Miscellaneous} captures names outside the classic \emph{enamax} type.
Apart from these there is \emph{timex}, which covers date \& time expressions and \emph{numex} which is for monetary values \& percent. 
Table \ref{table:NETypes} gives an overview of the different NE types.

\begin{table}[h!]
\scriptsize
\begin{tabular}{| l | l |}
\hline
\bf Named Entity Type & \bf Example \\
\hline
Person Names & Greta Garbo, John, Mary \\
Organisations& Benetton, Coca-Cola Gmbh\\
Locations&  Paris, Australia, Bristol\\
Miscellaneous& World Cup\\
 Date Expression& 01-02-2012, January, 6th, 1998 \\
Time Expression & 10 p.m.\\
Monetary Value &  \$15, 100 Euro   \\
Percent &   30\% \\
\hline
\end{tabular}
\caption{NE Examples}
\label{table:NETypes}
\end{table}

@@

\exg. \textbf{x:} The motor company ' Ford ' was founded and incorporated by Henry Ford .\\
      \textbf{y:} O O O O ORG  O  O O O O O PER PER O  \label{PredEx1b} \\

@@

Learning is implemented using the Structured Perceptron algorithm. To avoid overfitting, we average the parameters after the last iteration \cite{collins2002discriminative}. For finding the best sequence of labels, we use a modified version of the Viterbi algorithm. The algorithm finds the sequence $\hat{y}$, such that:

@@

\paragraph*{Node features}
Node features are features that only depend on the current token (node): the \emph{Token}, suffix and prefix, capitalisation.

@@

\begin{table}[h!]
\begin{tabular}{| l | l |}
\hline
\bf Feature & \bf Example \\
\hline
Token &  \textbf{``Amsterdam''}\\
Suffix& Amster\textbf{dam}\\
Prefix&  \textbf{San} Sebastian\\
Captitalized& \textbf{B}enetton\\
Number Pattern & \\
UPPERCASE &  \textbf{BENETTON}\\
POS-tag &  Benetton \textbf{NNP}   \\
Lemma &  produced $\Rightarrow$ produce \\
\hline
\end{tabular}
\caption{Node Features}
\label{table:node}
\end{table}

@@

\exg. Jack London went to New York .\\
      PER   PER   O    O  LOC LOC  O \\\label{seq1}

\exg. Jack London went to New York . \\ 
      PER  LOC    O    O  LOC LOC O \\\label{seq2}
